Small Language Modelling to Generate Playlists from Song Sequences

### 2023-12-12 docs/draft_report.tex
Draft Results

In contrast to the sequence-to-token models attempted with the LSTM architectures, sequence-to-sequence models are attempted with multi-head attention architectures. These models process the entire song sequence for a setlist into offset pairs to form the inputs and target outputs, where the original sequence is broken down into [0,...,n-1] tokens for the input sequence and [1,...,n] tokens for the output sequence. The sequences can be any size, although sequence lengths of 125 songs performed the best. The token sequences are encoded by an embedding layer composed of the sum of a learned N-dimensional vector embedding layer and a positional embedding layer. The positional embedding layer applies sine and cosine functions for fixed positional encodings. This is the same mechanism as is used in the original attention model [5].

@@

Summary of the Results.

\begin{table}
	\caption{Summary of Model Training Results}
	\label{results-table}
	\centering
	\begin{tabular}{lll}
		\toprule
		Model & Best Next-Song Validation Accuracy & Generated Sequence Quality \\
		\midrule
		2019 Model Recreation & $\sim$12.5\% & NA\footnote{Setlist generation was not tested for the 2019 model recreation attempts.} \\
		Model 1 & $\sim$18\% & Moderate \\
		Model 2 & 19.5\% & Poor \\
		Model 3 & 18.6\% & Good \\
		Model 4 & 15.9\% & Moderate \\
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{images/graph_attention_acc.png}
	\caption{Validation accuracies during training for the bi-directional LSTM model (model 2; \emph{left}) and for the self-attention model (model 3; \emph{right})}
\end{figure}

>>>>

In contrast to the sequence-to-token models attempted with the LSTM architectures, sequence-to-sequence models are attempted with multi-head attention architectures. These models process the entire song sequence for a setlist into offset pairs to form the inputs and target outputs, where the original sequence is broken down into [0, ..., $n-1$] tokens for the input sequence and [1, ..., $n$] tokens for the output sequence. The sequences can be any size, although sequence lengths of 125 songs performed the best. The token sequences are encoded by an embedding layer composed of the sum of a learned N-dimensional vector embedding layer and a positional embedding layer. The positional embedding layer applies sine and cosine functions for fixed positional encodings. This is the same mechanism as is used in the original attention model [5].

@@

The models were described in the previous section were trained for durations deemed appropriate based on stabilization of the validation accuracy for predicting the next song, and were tested for a variety of hyperparameter combinations. Training and Validation accuracy for Models 2 and 3 are shown in Figure 1, while Table 1 compares the best next song prediction validation accuracy for each model, and the results of the recreated 2019 model.

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{images/graph_attention_acc.png}
	\caption{Validation accuracies during training for the Bi-Directional LSTM model (Model 2; \emph{left}) and for the Self-Attention model (Model 3; \emph{right}).}
\end{figure}

\begin{table}
	\caption{Summary of Model Training Results}
	\label{results-table}
	\centering
	\begin{tabular}{lll}
		\toprule
		Model & Best Next-Song Validation Accuracy & Generated Sequence Quality \\
		\midrule
		2019 Model Recreation & $\sim$12.5\% & NA\footnote{Setlist generation was not tested for the 2019 model recreation attempts.} \\
		Model 1 & $\sim$18\% & Poor \\
		Model 2 & 19.5\% & Moderate \\
		Model 3 & 19.0\% & Moderate \\
		Model 4 & 16.8\% & Good \\
		\bottomrule
	\end{tabular}
\end{table}

All of the new models were able to outperform the recreated 2019 model in terms of next song accuracy, although the sequences that they generated were of varying degrees of quality. For example, Model 1 was able to predict a sequence of songs without repeats, but was not able to predict part-of-speech separators like 'set-1' or 'eos'. In contrast, Model 2 was able to predict part-of-speech separators but was unable to always do this in a way that made sense. It also missed many common sequential orderings of songs like 'The Horse' into 'Silent in the Morning'. Model 3 had better predictions for part-of-speech separators which were all sensical, although it repeated and alternated songs often. The band sometimes does this during shows so it is not an outlandish prediction, but Model 3 does this a little \emph{too} often to be considered a good prediction. Model 4 generated the best predictions with good placements for the part-of-speech separators and few repeated songs, despite it's lower next song accuracy compared to models 2 and 3.

Another evaluation method extends beyond predicting the $n^{th}$ song and instead assesses the model's capability to predict the next $n$ songs, or its accuracy in predicting the next sequence. To do this, Models 2, 3, and 4 were fed a seed sequence of a few songs from the beginning of a longer test setlist. Then the results of the model predictions from the seed sequence were compared to the actual next songs \emph{irrespective of song order}.

For each model, the setlist used for comparison was:

\begin{center}
	['farmhouse', 'first-tube', 'twist', 'divided-sky', 'ginseng-sullivan', 'carini', 'whats-the-use', 'wildcard', 'set-2', 'down-with-disease', 'the-moma-dance', 'piper', 'fee', 'gotta-jibboo', 'saw-it-again', 'split-open-and-melt', 'cavern', 'david-bowie', 'set-e', 'the-squirming-coil', 'eos']
\end{center}

Table 2 shows the results of this exercise. Models 2, 3, and 4 were all able to achieve similar accuracies in this regard. They all correctly predicted part-of-speech separators which would be used, and were able to predict a few songs which would appear in the next setlist.

\begin{table}
	\caption{Summary of Model Training Results}
	\label{generation-table}
	\centering
	\begin{tabular}{lll}
		\toprule
		Model & Common Songs (Original \& Prediction) & Accuracy \\
		\midrule
		Model 2 & ['set-2', 'down-with-disease', 'set-e', 'piper', 'eos', 'the-moma-dance'] & 33.33\% \\
		Model 3 & ['wildcard', 'set-2', 'set-e', 'david-bowie', 'cavern', 'eos'] & 33.33\% \\
		Model 4 & ['set-e', 'wildcard', 'piper', 'the-moma-dance', 'eos', 'set-2'] & 33.33\% \\
		\bottomrule
	\end{tabular}
\end{table}
